---
title: "Homework 02: Explore Gapminder and use dplyr"
author: "Ke Dai"
date: "2016/9/27"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bring rectangular data in

```{r}
library(gapminder)
library(tidyverse)
```

## Smell test the data

**Explore the gapminder object**

- **Is it a data.frame, a matrix, a vector, a list?**
```{r}
is.data.frame(gapminder)
is.matrix(gapminder)
is.vector(gapminder)
is.list(gapminder)
```
Gapminder is a data.frame and also a list.

- **What’s its class?**
```{r}
class(gapminder)
```
The class of gapminder is table or data.frame.

- **How many variables/columns?**
```{r}
ncol(gapminder)
```
Gapminder has 6 variables or columns.

- **How many rows/observations?**
```{r}
nrow(gapminder)
```
Gapminder has 1704 observations or rows.

- **Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?**
```{r}
str(gapminder)
glimpse(gapminder)
dim(gapminder)
```
We can also use str, glimpse and dim function to get the size. Compared to nrow and ncol, dim can get the number of both row and column. And str and glimpse can get the name and type of each variable in addition to the number of row and column. So str and glimpse seem more powerful.

- **What flavor is each variable?**

We can get the flavor of each variable using str(gapminder) as mentioned before. Country and continent variable are of factor flavor. Year, lifeExp, pop and gdpPercap variable are of numeric flavor.

## Explore individual variables
**Pick at least one categorical variable and at least one quantitative variable to explore.**

- **Characterize what’s possible, i.e. all possible values or max vs. min … whatever’s appropriate.**
- **What’s typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.**
- **Feel free to use summary stats, tables, figures. We’re NOT expecting high production value (yet).**
```{r}
levels(gapminder$continent)
```
For categorical variable, I choose continent. We can see that the possible values of continent are Africa, America, Asia, Europe, and Oceania respectively. 

```{r}
summary(gapminder$lifeExp)
```
For quantitative variable, I choose lifeExp. We can see that the minimum value is 23.60. The maximum value is 82.60. The median value is 60.71. The mean value is 59.47. The first quantile is 48.20. The third quantile is 70.85.

```{r}
ggplot(data = gapminder, mapping = aes(x = as.integer(lifeExp))) + 
  geom_bar()
range(gapminder$lifeExp)
sd(gapminder$lifeExp)
```

From the figures above, we can see that the most typical value of lifeExp is around 70 and the values below 30 are odd. We can also get the value range of lifeExp (from 23.599 to 82.603) using range function instead of min and max function or summary function. The distribution of lifeExp is well demonstrated by the bar graph.

## Explore various plot types
**See the ggplot2 tutorial, which also uses the gapminder data, for ideas.**

**Make a few plots, probably of the same variable you chose to characterize numerically. Try to explore more than one plot type. Just as an example of what I mean:**

- **A scatterplot of two quantitative variables.**

- **A plot of one quantitative variable. Maybe a histogram or densityplot or frequency polygon.**

- **A plot of one quantitative variable and one categorical. Maybe boxplots for several continents or countries.**

**You don’t have to use all the data in every plot! It’s fine to filter down to one country or small handful of countries.**

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) +
  scale_x_log10() +
  geom_point() +
  facet_wrap(~ continent)
```
```{r}
library(mosaic)
library(lattice)
histogram(gapminder$lifeExp)
freqpolygon(gapminder$lifeExp)
densityplot(gapminder$lifeExp)

ggplot(gapminder, mapping = aes(x = lifeExp)) +
  geom_histogram(fill = "red")
ggplot(gapminder, mapping = aes(x = lifeExp)) +
  geom_freqpoly(color = "blue")
ggplot(gapminder, mapping = aes(x = lifeExp)) +
  geom_density(color = "green")
```
```{r}
ggplot(data = gapminder, mapping = aes(x = continent, y = lifeExp)) +
  geom_boxplot()
```

As shown above, I  plot a scatterplot of lifeExp and gdpPercap, 2 histograms, 2 densityplots and 2 frequency polygons of life Exp using different libraries, and a plot of lifeExp and continent.

## Use filter(), select(), and %>%
**Use filter() to create data subsets that you want to plot.**
**Practice piping together filter() and select(). Possibly even piping into ggplot().**
```{r}
gapminder %>% 
   filter(country == "Canada" | country == "China") %>% 
   select(country, year, lifeExp) %>%
   ggplot(aes(x = year, y = lifeExp, fill = country)) +
   geom_bar(stat = "identity", position = "dodge")

gapminder %>% 
   filter(country == "Canada" | country == "China") %>% 
   select(country, year, pop) %>%
   ggplot(aes(x = year, y = pop, color = country)) +
   scale_y_log10() +
   geom_point()

gapminder %>% 
   filter(country == "Canada" | country == "China") %>% 
   select(country, year, gdpPercap) %>%
   ggplot(aes(x = year, y = gdpPercap, color = country)) +
   scale_y_log10() +
   geom_point() +
   geom_path()
```

As I am from China, I am interested in make some comparison between China and Canana. In the first graph, I compared the lifeExp of the two countries. We can see that the present life expectancy of China is still shorter than that of Canada although it has grown more rapidly in the last fifty years. In the second graph, I compared the pop of them. We can see that the pupulation of China has more than doubled while the population of Canada remains stable. In the third graph, I compared the gdpPercap between them. We can see that the GDP per capita of both countries has experienced a significant growth.

## But I want to do more!
**For people who want to take things further.**

**Evaluate this code and describe the result. Presumably the analyst’s intent was to get the data for Rwanda and Afghanistan. Did she succeed? Why or why not? If not, what is the correct way to do this?**
```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```
The code above is not correct, it misses 12 obsersations in total. The correct code is as follows:
```{r}
filter(gapminder, country %in% c("Rwanda", "Afghanistan"))
```

**Present numerical tables in a more attractive form, such as using knitr::kable().**

```{r}
gapminder %>% 
  filter(country == "Canada" | country == "China") %>% 
  select(country, year, lifeExp, pop, gdpPercap) %>%
  knitr::kable()
```

**Use more of the dplyr functions for operating on a single table.**
```{r}
## using *arrange()* and *desc()* to change the order of the subset of gapminder
gapminder %>%
  filter(country == "Canada" | country == "China") %>% 
  select(country, year, lifeExp, pop, gdpPercap) %>% 
  arrange(desc(year))

## using *mutate()* to add a new variable gdp to the gapminder data set.
mutate(gapminder, gdp = gdpPercap * pop)

## using *group_by()*, *summarise()* and ggplot() to plot the average life expectancy of each continent from 1952 to 2007.
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(avgLifeExp = mean(lifeExp, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = avgLifeExp, color = continent)) +
  geom_point() +
  geom_path()
```
  
## Report your process
Before starting doing Homework 02, I first spent serveral hours reading Explore part of the book ***R for Data Science***. For the first time reading, I did not force myself to memorize the syntax of each command exactly. I just want to have an impression of the content. And then I get started with the homework. When I got stuck on a specific question, I tried to find the answer or get some ideas from the book or the notes in the lecture. I have to say this book is really useful. I can find the answers to most of my questions in it. For the questions I cannot solve through the book, I tried to search the answers or relative information by google or baidu (a popular search engine for Chinese people). Although I have spared no effort to figure out every question I had, there are still some questions remaining to be solved. Anyway, the process is interesting. I learned a lot from this assignment.

## My questions
1. What is the difference among *typeof()*, *mode()* and *class()*?
2. When I apply *is.data.frame()* and *is.list()* to gapminder, both of them return true. So is it that *is.<type>(obj)* returns true as long as this object is of the given type or class.
3. In the below code chunk, I create a vector. Why do *typeof(a)* and *mode(a)* not return vector?
```{r}
a = c(1, 2, 3)
typeof(a)
mode(a)
```
