---
title: 'Homework 10: Get data from the web'
author: "Ke Dai"
date: "2016年12月6日"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data and tools
```{r}
library(gapminder)
library(countrycode)
library(geonames)
library(ggplot2)
library(tidyverse)
library(stringr)
library(knitr)
library(httr)
library(purrr)
options(geonamesUsername = "ke_dai")
```

## Make API queries “by hand” using httr
Get 7 day weather forecast for Vancouver from OpenWeatherMap.
```{r}
## Get the raw data
response <- GET('api.openweathermap.org/data/2.5/forecast/daily?q=Vancouver&units=metric&cnt=7&APPID=2aa01b3164554605bbc8a44a976ae76a')
status_code(response)
headers(response)[["content-type"]]
rawData <- content(response, type = "application/json")
str(rawData, max.level = 3)

## Transfrom the raw data into a data frame
weatherVancouver <- tibble(
  temp = rawData$list %>% map("temp"),
  pressure = rawData$list %>% map_chr("pressure"),
  humidity = rawData$list %>% map_chr("humidity"),
  windSpeed = rawData$list %>% map_chr("speed"),
  windDeg = rawData$list %>% map_chr("deg")
)

kable(weatherVancouver)

## Turn the nested data frame into a regular one
weatherVancouver <- weatherVancouver %>% 
  mutate(
    day = map(temp, "day"),
    night = map(temp, "night")
  ) %>% 
  select(-temp) %>% 
  unnest()

kable(weatherVancouver)

## create a date tibble
date <- tibble(date = str_c("12.", 16:22))
date

## combine date with weatherVancouver
weatherVancouver <- cbind(date, weatherVancouver)
kable(weatherVancouver)

## gather weatherVanvouer
weatherVancouver <- weatherVancouver %>% 
  gather(`day`, `night`, key = "duration", value = "temp")

kable(weatherVancouver)

## plot the variation of day and night temperature 
weatherVancouver %>% 
  ggplot(aes(x = date, y = temp, color = duration)) +
  geom_point() +
  labs(title = "The day and night temperature forecast for Vancouver in the next 7 days", x = "Date", y = "Temperature")

## plot the variation of pressure
weatherVancouver %>% 
  ggplot(aes(x = date, y = pressure)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "The pressure forecast for Vancouver in the next 7 days", x = "Date", y = "Pressure")
```


## Use an R package that wraps an API
Combine gapminder and data from geonames. Install the geonames package (on CRAN, on GitHub). Make a user account and use geonames to access data about the world’s countries. Use data from geonames and gapminder to investigate the population density against time.
```{r}
## change country variable in gapminder to countryCode variable
gapminderCode <- gapminder %>% 
  mutate(countryCode = countrycode(country, "country.name", "iso2c"))

## get countryInfo data
countryInfo <- GNcountryInfo()

## select countryCode and areaInSqKm variables from countryInfo
countryArea <- countryInfo %>% 
  select(countryCode, area = areaInSqKm) %>% 
  mutate(area = as.double(area))

kable(countryArea[1:20, ])

## left join gapminderCode with countryPos
gapminderArea <- gapminderCode %>% 
  left_join(countryArea, by = "countryCode")

kable(gapminderArea[1:20, ])

## calculate the population density
gapminderPopDensity <- gapminderArea %>% 
  mutate(popDensity = pop/area) %>% 
  select(-area)

kable(gapminderPopDensity[1:20, ])
 
## plot the graph of population density agianst time
ggplot(subset(gapminderPopDensity, continent != "Oceania"),
       aes(x = year, y = popDensity, group = country, color = country)) +
  geom_line(lwd = 1, show.legend = FALSE) + facet_wrap(~ continent) +
  scale_color_manual(values = country_colors) + theme_bw() +
  theme(strip.text = element_text(size = rel(1.1))) + scale_y_log10()
```

## Report my process
For this homework, I explored two ways to get data from the internet into R: 1. Make API queries “by hand” using httr; 2. Use an R package that wraps an API. The second way is not very difficult for me as I have tried combining gapminder and data from geonames in Homework 04. However, I spent too much time exploring the first approach. I still have some problems with processing the nested list although I finished my task successfully. I will continue to learn more skills about R in the future. Congratuations! I have finished all of ten assignments. I learned a lot from this course. I appreciate Jennifer's and all the TAs' effort on this course. Finally, hope everyone have a good vacation.
