---
title: 'Homework 10: Get data from the web'
author: "Ke Dai"
date: "2016年12月6日"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data and tools
```{r}
library(gapminder)
library(countrycode)
library(geonames)
library(ggplot2)
library(dplyr)
library(knitr)
options(geonamesUsername = "ke_dai")
```

## Use an R package that wraps an API
Combine gapminder and data from geonames. Install the geonames package (on CRAN, on GitHub). Make a user account and use geonames to access data about the world’s countries. Use data from geonames and gapminder to investigate the population density against time.
```{r}
## change country variable in gapminder to countryCode variable
gapminderCode <- gapminder %>% 
  mutate(countryCode = countrycode(country, "country.name", "iso2c"))

## get countryInfo data
countryInfo <- GNcountryInfo()

## select countryCode and areaInSqKm variables from countryInfo
countryArea <- countryInfo %>% 
  select(countryCode, area = areaInSqKm) %>% 
  mutate(area = as.double(area))

kable(countryArea[1:20, ])

## left join gapminderCode with countryPos
gapminderArea <- gapminderCode %>% 
  left_join(countryArea, by = "countryCode")

kable(gapminderArea[1:20, ])

## calculate the population density
gapminderPopDensity <- gapminderArea %>% 
  mutate(popDensity = pop/area) %>% 
  select(-area)

kable(gapminderPopDensity[1:20, ])
 
## plot the graph of population density agianst time
ggplot(subset(gapminderPopDensity, continent != "Oceania"),
       aes(x = year, y = popDensity, group = country, color = country)) +
  geom_line(lwd = 1, show.legend = FALSE) + facet_wrap(~ continent) +
  scale_color_manual(values = country_colors) + theme_bw() +
  theme(strip.text = element_text(size = rel(1.1))) + scale_y_log10()
```

## Report my process
For this homework, I explored getting data from the internet into R using an R package that wraps an API. It's not very difficult for me as I have tried combining gapminder and data from geonames in Homework 04. Congratuations! I have finished all of ten assignments. I learned a lot from this course. I appreciate Jennifer's and all the TAs' effort on this course. Finally, hope everyone have a good vacation.
